Structured Versus Decorated Cospans

### 2020-12-04 cospans.tex
worked on Conclusions

% Structured and decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2020/12/3

@@

We have given conditions under which a decorated cospan double category is isomorphic to a structured cospan double category, in \cref{thm:equiv}. The converse question is also interesting: is every structured cospan double category equivalent to a decorated cospan double category? The answer is similar to the previous one: affirmative, under certain conditions that allow us to make the passage from an appropriate functor $L\colon \A\to \X$ to an appropriate pseudofunctor $F\colon \A\to\Cat$, namely in the reverse direction of the earlier \cref{eq:FtoL}.

Let us now give a sketch of the relevant story. Suppose $\A$ and $\X$ have finite colimits and $L\maps\A\to\X$ is a finite-colimit preserving functor -- which by \cref{thm:SC} are the required conditions for constructing the symmetric monoidal double category of structured cospans. If $L$ has a right adjoint `left inverse' (meaning the unit is the identity) $U\maps\X\to\A$, which moreover strictly preserves the choice of pushouts in $\A$ and $\X$, it can be shown\footnote{This will appear in the forthcoming paper \cite{CV}.} that $U$ is an opfibration. Consequently, $U$ corresponds to a pseudofunctor $F\colon \A\to\Cat$ by the reverse of the Grothendieck construction, i.e. the second part of \cref{thm:Grothendieck}. Furthermore, if $U$ preserves finite coproducts, by the special case of the cocartesian monoidal Grothendieck construction discussed under \cref{lem:MonGroth}, the corresponding pseudofunctor $F$ has a lax monoidal structure $(\A,+)\to(\Cat,\times)$. As a result, by \cref{DC}, $F$ has now enough structure to induce the symmetric monoidal double category of decorated cospans $F\lCsp$, and further it can then be verified that it is isomorphic to the one of structured cospans.

{\chris Give some sort of excuse why we don't do the full result here?? Because of missing - for us now - steps? Or other things?}

To give a better sense of how this lax monoidal pseudofunctor $F\maps\A\to\Cat$ is constructed, $F(a)$ is defined to be the pre-image of $a$ under $U$, namely the category of all objects in $x\in\X$ such that $U(x)=a$ and morphisms $k\maps x\to y\in\X$ such that $U(k)=1_a$. Given a morphism $f\maps a\to b$, there is a functor $F(f)\maps F(a)\to F(b)$ that maps some $x$ to the following pushout
\begin{displaymath}
 \begin{tikzcd}
La\ar[r,"Lf"]\ar[d,"\varepsilon_x"']\ar[dr,phantom,near end,"\ulcorner"] & Lb\ar[d] & \\
x\ar[d,dotted]\ar[r] & x+_{La}Lb\ar[d,dotted] & \textrm{in }\X \\
a\ar[r,"f"'] & b & \textrm{in }\A
 \end{tikzcd}
\end{displaymath}
where $\varepsilon_x\colon LU(x)=L(a)\to x$ is the counit of $L\dashv U$. The fact that $U$ strictly preserves pushouts is necessary to show that the pushout is mapped, via $U$, directly down to $b$.

>>>>

% Structured and decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2020/12/4

@@

We have given conditions under which a decorated cospan double category is isomorphic to a structured cospan double category, in \cref{thm:equiv}. The converse question is also interesting: is every structured cospan double category equivalent to a decorated cospan double category? The answer is similar to the previous one: yes, under certain conditions that let us pass from an appropriate functor $L \maps \A \to \X$ to an appropriate pseudofunctor $F \maps \A \to \Cat$.

Let us now sketch the story; details will appear in a forthcoming paper \cite{CV}.  Suppose the conditions hold for constructing the double category of structured cospans ${}_L \lCsp(\X)$ as in \cref{thm:SC}.  That is, suppose $\A$ and $\X$ have finite colimits and $L \maps \A \to \X$ preserves them.    If $L$ also has a right adjoint `left inverse' (meaning the unit is the identity) $U \maps \X \to \A$, which moreover strictly preserves the chosen pushouts, it can be shown that $U$ is an opfibration.  Consequently, $U$ corresponds to a pseudofunctor $F \maps \A \to \Cat$ by the inverse Grothendieck construction, as in the second part of \cref{thm:Grothendieck}. Furthermore, if $U$ preserves finite coproducts, $F$ acquires the structure of a lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat,\times)$ by the special case of the cocartesian monoidal Grothendieck construction discussed under \cref{lem:MonGroth}.  As a result, $F$ has now enough structure to induce a double category of decorated cospans $F\lCsp$  as in \cref{DC}.  Finally, it can be shown that the structured and decorated cospan double categories categories are equivalent as symmetric monoidal double categories: ${}_L \lCsp(\X) \simeq F\Csp$.

To give a better sense of how the pseudofunctor $F \maps \A \to\Cat$ is constructed: for each object $a \in A$, $F(a)$ is defined to be the fiber of $U$ over $a$, namely the category of all objects in $x \in \X$ such that $U(x)=a$ and morphisms $k\maps x\to y$ such that $U(k)=1_a$.  Given a morphism $f \maps a \to b$, there is a functor $F(f) \maps F(a)\to F(b)$ that maps $x \in F(a)$ to the following pushout:
\begin{displaymath}
 \begin{tikzcd}
La\ar[r,"Lf"]\ar[d,"\varepsilon_x"']\ar[dr,phantom,near end,"\ulcorner"] & Lb\ar[d] & \\
x\ar[d,dotted]\ar[r] & x+_{La}Lb\ar[d,dotted] & \textrm{in }\X \\
a\ar[r,"f"'] & b & \textrm{in }\A
 \end{tikzcd}
\end{displaymath}
where $\varepsilon_x\colon LU(x)=L(a)\to x$ is the counit of the adjunction $L\dashv U$. The fact that $U$ strictly preserves pushouts is necessary to show that the pushout is mapped, via $U$, directly down to $b$.
